Concurrent Programming with CXL

### 2020-06-24 book/paper.tex
Made pass over actor chapter

Some programming languages favor a different way of implementing
synchronization using so-called \emph{actors}~\cite{HBS73}.  Actors are
processes that have only private memory and communicate through message passing.
See Figure~\ref{fig:actorpic} for an illustration.
Given that there is no shared memory in the actor model (other than the message
queues, which have built-in synchronization), there is no need
for critical sections.  Instead, some sequential process owns a particular
piece of data and other processes access it by sending request messages
to the process and optionally waiting for response messages.  Each process
handles one message at a time, serializing all access to the data it owns.

@@

Figure~\ref{fig:actor} shows a CXL solution to the producer/consumer problem
(with an unbounded buffer)
using an actor (as illustrated in Figure~\ref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \texttt{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \texttt{balance} and \texttt{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

The \texttt{produce} method takes two arguments: a pointer to the queue
of the \texttt{pc\_actor} and the item.  The \texttt{consume} method
takes two queue pointer arguments.  The first is a pointer to the queue
of the \texttt{pc\_actor}.  The second is a pointer to the queue of the
invoker: it is the queue to which the response must be posted.

@@

>>>>

Some programming languages favor a different way of implementing
synchronization using so-called \emph{actors}~\cite{HBS73}.  Actors are
processes that have only private memory and communicate through
\emph{message passing}.
See Figure~\ref{fig:actorpic} for an illustration.
Given that there is no shared memory in the actor model (other than the message
queues, which have built-in synchronization), there is no need
for critical sections.  Instead, some sequential process owns a particular
piece of data and other processes access it by sending request messages
to the process and optionally waiting for response messages.  Each process
handles one message at a time, serializing all access to the data it owns.
As message queues are FCFS, starvation is prevented.

@@

Figure~\ref{fig:actor} shows a CXL solution to the
``multiple producer / multiple consumer problem'' (with an unbounded buffer)
using an actor (as illustrated in Figure~\ref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \texttt{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \texttt{balance} and \texttt{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

The shared memory consists entirely of message queues.
The \texttt{produce} method takes two arguments: a pointer to the queue
of the \texttt{pc\_actor} and the item.  The \texttt{consume} method
takes two queue pointer arguments.  The first is a pointer to the queue
of the \texttt{pc\_actor}.  The second is a pointer to the queue of the
invoker: it is the queue to which the response must be posted.

@@

TODO: Talk about seqlock perhaps.
